\documentclass[11pt,a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[english]{babel}
\usepackage{geometry}
\usepackage{longtable}
\usepackage{booktabs}
\geometry{margin=2.5cm}

\title{System Architecture Description\\AymaraLang}
\author{AymaraLang Project}
\date{\today}

\begin{document}
\maketitle

\section{Overview}
AymaraLang is composed of a modular compiler (\texttt{aymc}), a minimal runtime, and supporting tools. The architecture follows the classic compiler pipeline with decoupled modules.

\section{Functional architecture}
\begin{longtable}{@{}p{3cm}p{4cm}p{4cm}p{4cm}@{}}
\toprule
\textbf{Module} & \textbf{Responsibility} & \textbf{Inputs} & \textbf{Outputs} \\ \midrule
Lexer & Source code tokenization & \texttt{.aym} & Tokens \\
Parser & AST construction & Tokens & AST \\
Semantic & Semantic and type analysis & AST & Validated AST + symbols \\
Codegen & NASM/LLVM generation & Validated AST & \texttt{.asm}/\texttt{.ll} \\
Linker & Assembly/linking & \texttt{.asm} & Binary \\
\bottomrule
\end{longtable}

\section{Physical architecture (deployment)}
\begin{itemize}
    \item \textbf{Local environment:} Linux/Windows.
    \item \textbf{Dependencies:} NASM, GCC/LD or MinGW, C++17 compiler.
    \item \textbf{Artifacts:} \texttt{.asm}, \texttt{.ll}, native binary, error logs.
\end{itemize}

\section{Key interfaces}
\begin{itemize}
    \item \textbf{CLI:} \texttt{aymc [options] file.aym}.
    \item \textbf{File system:} source reads, artifact writes.
    \item \textbf{Runtime:} basic I/O functions used by the executable.
\end{itemize}

\section{Data and information}
\begin{itemize}
    \item \textbf{Tokens:} type, text, line, and column.
    \item \textbf{AST:} expression and statement nodes.
    \item \textbf{Symbols:} scope tables for variables and functions.
\end{itemize}

\section{Design constraints}
\begin{itemize}
    \item C++17 as the baseline standard.
    \item x86\_64 NASM codegen as the primary backend.
    \item LLVM as an experimental backend.
\end{itemize}

\end{document}
